Generalized Tractability

### 2023-05-12 GeneralizedTractability_230501.tex
Updates from Overleaf

The question of tractability is one of determining how fast the complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding.

@@

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= \max\{1,\varepsilon^{-p}\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0.
\]
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (1,\min\{\lambda_{i,d}^p\}) < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NW08}.
\end{example}

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,(p,q))= (\max\{1,\varepsilon^{-p}\})\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p,q \ge 0.
\]
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_{p,q} d^{q}}^\infty (\min\{1,\lambda_{i,d}^p\}) d^{-q} < \infty \text{ for some } p,q \ge 0, L_{p,q} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08} \fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NW08}}.
\end{example}

\begin{example}[Strong exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= \left[\max\{1,\log(1+\varepsilon^{-1}\})\right]^p.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L_{\vp} d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p  < \infty.
\]
\end{example}
\begin{example}[Exponential polynomial tractability]

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,(p,q))=\left[\max\{1,\log(1+\varepsilon^{-1}\})\right]^p\, d^q.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L_{\vp} d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p d^{-q} < \infty.
\]
\end{example}
\bigskip

\begin{example}[Algebraic quasi-polynomial tractability]
\begin{definition}
    As defined in \cite{GW11}, we say that a problem is \emph{ quasi-polynomially tractable (QPT)} if there exist two constants $C$, $t > 0$, such that
\begin{equation}\label{expqpt}
    n(\varepsilon,d) \leq C\exp\{t(1+\log(\varepsilon^{-1}))(1+\log(d))\}
\end{equation}
for all $\varepsilon \in (0,1]$ and $d\in\mathbb{N}$.\\
\end{definition}
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\})(1+\log d)\}
\]
Observe that this is a special case of the $T$-tractability function, where the function is polynomial of exponent $1+\log d$ and $1+\log\max\{\varepsilon^{-1},1\}$.

Then Theorem \ref{sec:tractability} yields the iff-condition as follows

\[
\sup_{d\in\mathbb{N}}\sum_{i=Lpd^{q_1}} \left(\min\{\lambda_{i,d},1\} \right)^{p(1+\ln d)}d^{-q} < \infty
\]
\end{example}

\kachi{In the following we introduce two specific cases of tractability which depend on the form of the function $T$.
\begin{example}[(Strong) Separable tractability]
Let the tractability function $T$,
be defined by
\[
 T(\varepsilon^{-1},d,(p,q))= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p,q \ge 0.
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$. In addition, we assume the following conditions as in (\ref{eq:ptauassume}) 
%\begin{align*}
 %(\phi(\varepsilon^{-1},p))^{\tau_1} &\le K_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)\\ (\psi(d,q))^{\tau_2} &\le K_{q,\tau_2} \psi (d,\tau_2 q)\\\end{align*}
 \begin{align*}
     [T(\varepsilon^{-1},d,(p,q))]^\tau \leq K_{(p,q),\tau}T(\varepsilon^{-1},d,(\tau p,\tau q))
 \end{align*}
where $\tau>0$.
Observe that for this $T$, we have strong tractability if the following condition is satisfied
\[
 T(\varepsilon^{-1},1,p)= \phi(\varepsilon^{-1},p)\, \psi(1,0) = \phi(\varepsilon^{-1},p)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0.
\]
Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent. We say that a problem is strongly separable tractable iff there exists $p>0$ and an integer $C(p)\ge 0$ such that
\begin{equation} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L(p,1)}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation}

%As the name implies, the roles of $\varepsilon^{-1}$ and $d$ are separated. 
\end{example}
\begin{example}[Non-separable symmetric tractability]
    Let the tractability function $T$, be defined by
\[
 T(\varepsilon^{-1},d,(p,q))= \exp\{\phi(\varepsilon^{-1},p)\, \psi(d,q)\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p\ge 0.
\]  with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$.
\end{example}
}

@@

\section{Try Restricted Region}

Suppose that the domain of interesting $(\varepsilon^{-1},d)$ is given by 
\begin{equation}
    \Omega  = \{(\varepsilon^{-1},d) : \varepsilon > E(d) \ge 0, \ d \in \naturals \},
\end{equation}
where $E$ is a non-increasing function.  Furthermore, for all $d \in \naturals$, let $N(d) = \min \{n : \lambda_{n+1,d} \le E(d) \}$.  This means that 
\[
\lambda_{1,d}^{-1} \le \cdots \le \lambda_{N(d),d}^{-1} < E(d) \le \lambda_{N(d)+1,d}^{-1}, \qquad (\lambda_{1,d}^{-1},d), \ldots, (\lambda_{N(d),d}^{-1},d) \in \Omega.
\]

\begin{theorem}\label{thm_main_strong_tract3}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is strongly $T$-tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L_{\vp} > 0$ such that
\begin{equation} \label{eq:strong_tractiff3}
     S_{\vp}:=\sup_{d \in \naturals} \sum_{i = L_{\vp}}^{N(d)} \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff3} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tractiff3} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{strct}} = \widetilde{\cp}_{\textup{strct}}$, and the set of optimal strong tractability parameters is
\[
	\cp_{\textup{sopt}} =
	\{\vp^* \in \widetilde{\cp}_{\textup{strct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{strct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

\begin{proof}
\textbf{Sufficient condition:}\newline
We make the first part of the argument in some generality so that it can be reused in the proof of Theorem \ref{thm_main_tract2} for tractability.

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},d',\vp)$ are non-decreasing in $i$ for any $d'\in\{1, \ldots, d\}$. In particular, for all $n\in \{1, \ldots, N(d')\}$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{1,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},d',\vp)} \le \cdots \le \frac{1}{T(\lambda_{1, d}^{-1},d',\vp)} \\
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp) }
    \le \frac 1n \sum_{i=1}^n  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) }
    \le \frac 1n \sum_{i=1}^{N(d')}  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n \ge T(\varepsilon^{-1},d',\vp) \sum_{i=1}^{N(d')} \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le
   \frac 1n \sum_{i=1}^{N(d')} \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) } \le \frac{1}{T(\varepsilon^{-1},d',\vp)} \qquad \forall \varepsilon > 0, \ d,d' \in \naturals.
\end{multline}

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
    \begin{align} %\label{eq:comp_equiv}
    \nonumber
    \comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
    \nonumber
    & = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},d',\vp)}\le \frac{1}{T(\varepsilon^{-1},d',\vp) }\biggr\} \qquad \forall d' \in \naturals\\
    \nonumber
    & \le  \min \biggl\{n \in \natzero : n \ge T(\varepsilon^{-1},d',\vp) \sum_{i=1}^{N(d')} \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \biggr\} \quad \forall d' \in \naturals \qquad \text{by \eqref{eq:implicate_a}} \\
    & \le  T(\varepsilon^{-1},d',\vp) \sum_{i=1}^{N(d')} \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} + 1 \quad \forall d' \in \naturals. \label{eq:compUP}
\end{align}

Now we take this upper bound further, specializing to the case of $d'=1$:
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},1, \vp) \Biggl [ \sum_{i=1}^{L_{\vp}-1} \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}
       + \underbrace{\sum_{i=L_{\vp}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{\le S_{\vp} <\infty\ \text{by \eqref{eq:strong_tractiff3}} }
       + \frac{1}{T(\varepsilon^{-1},1, \vp)} \Biggr] \\
       & \le T(\varepsilon^{-1},1,\vp) \underbrace{\left[\frac{(L_{\vp}-1)}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp} + \frac{1}{T(0,1, \vp)}\right]}_{=:C_{\vp}}\\
       & =  C_{\vp}T(\varepsilon^{-1},1,\vp).
%       & \le \underbrace{2 C_{1,\vp}}_{\text{since } T(\varepsilon^{-1},1,\vp) \ge 1} T(\varepsilon^{-1},1,\vp)
%       = \underbrace{C_{\vp}}_{:=2 C_{1,\vp}}T(\varepsilon^{-1},1,\vp).
\end{align*}
This means that we have strong $T$-tractability via \eqref{eq:strong_tract_def}, and verifies the sufficiency of \eqref{eq:strong_tractiff3}.

\bigskip
\noindent \textbf{Necessary condition:} \\
Suppose that we have strong
$T$-tractability as defined in \eqref{eq:strong_tract_def}. That is, for some $\vp \ge \vzero$ there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)
\qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\]
Since the sequence of eigenvalues $\lambda_{1,d}, \lambda_{2,d}, \ldots $ is non-increasing, we have
\begin{equation}\label{eq:lambda_K_strong3}
\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1,d}\le \varepsilon.
\end{equation}

Define
\[
i(\varepsilon,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1, \quad
i(\infty,\vp) = \lim_{\varepsilon \to \infty}  i(\varepsilon,\vp) \ge 1 .
\]
Thus, it follows by \eqref{eq:lambda_K_strong3} that $\lambda_{i(\varepsilon,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
i(\varepsilon,\vp)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1 \le C_{\vp} T(\lambda_{i(\varepsilon,\vp),d}^{-1},1,\vp)+1, \qquad \forall \varepsilon > 0,
\]
since
$T(\cdot,1, \vp)$ is non-decreasing.

>>>>

The question of tractability is one of determining how fast the complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding. \kachi{In this paper, we investigate such a function that we refer to as the generalized tractability function, denoted by $T$. A proper definition is given in Section \ref{sec:spt}. However, we define a tractability function
\[
 T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\]
which depends on $\varepsilon^{-1}$, $d$ and the parameter $\vp$ respectively.
}

@@

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]  where in this case $\vp$ is a scalar.
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (\min\{1,\lambda_{i,d}^p\}) < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NW08}.
\end{example}

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \max\{1,\varepsilon^{-p}\}\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where in this case $\vp = (p,q)$.
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability:
\[
 \sup_{d\in\naturals} \sum_{i=L_{(p,q)} d^{q}}^\infty (\min\{1,\lambda_{i,d}^p\}) d^{-q} < \infty \text{ for some } p,q \ge 0, L_{p,q} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08} \fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NW08}}.
\end{example}

\begin{example}[Strong exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^p, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
where $\vp$ is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for strong exponential polynomial tractability \[
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty (\min\{1,\log(1+\lambda_{i,d}\})^{p}  < \infty.
\] In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1}\})\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability
\[
 \sup_{d\in\naturals} \sum_{i=L_{(p,q)} d^{q}}^\infty [\min\{1,\log(1+\lambda_{i,d}\})]^p d^{-q} < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
\end{example}

\begin{example}[Algebraic quasi-polynomial tractability] Here we consider another common notions of tractability.

As defined in \cite{GW11}, we say that a problem is \emph{quasi-polynomially tractable (QPT)} if there exist two constants $C$, $t > 0$, such that
\begin{equation}\label{expqpt}
    \comp(\varepsilon,d) \leq C\exp\{t(1+\log(\varepsilon^{-1}))(1+\log(d))\}
\end{equation}
for all $\varepsilon \in (0,1]$ and $d\in\mathbb{N}$.\\

\noindent Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\}))(1+\log d)\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\varepsilon^{-1}\}]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\varepsilon^{-1}\}$ and $d$ which are, polynomial or not exponential in powers of $\max\{1,\varepsilon^{-1}\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the algebraic case
\[
\sup_{d\in\mathbb{N}}\sum_{i=Lp\,e^pd^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)}e^{-p}d^{-p} < \infty \qquad \text{for some} p \geq 0, L_p \in \naturals.
\]
\end{example}
\begin{example}[Exponential quasi-polynomial tractability]

\noindent Suppose that the tractability function, $T$, is defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log d)\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\log(1+\varepsilon^{-1}\})]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\log(1+\varepsilon^{-1})\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the exponential case
\[
\sup_{d\in\mathbb{N}}\sum_{i=Lp\,e^pd^{p}} \left(\min\{1,\log(1+\lambda_{i,d})\} \right)^{p(1+\ln d)}e^{-p}d^{-p} < \infty
\]
\end{example}

\kachi{In the following we introduce two specific cases of tractability which depend on the form of the function $T$.
\begin{example}[(Strong) Separable tractability]
Suppose that
\[
 T(\varepsilon^{-1},d,\vp)= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty).
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ and where $\vp = (p,q)$.

%In addition, we assume the following conditions as in (\ref{eq:ptauassume}). That is, there exists a $K_{(p,q),\tau}$ depending on $p,q$ and $\tau$, but independent of $\varepsilon$ and $d$, such that
%\begin{equation} \label{eq:tractiff}    \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = C(p,q)}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p))} < \infty.\end{equation}

 %\begin{align*}
  %   [T(\varepsilon^{-1},d,\vp]^\tau \leq K_{\vp,\tau}T(\varepsilon^{-1},d,\tau\vp) \quad \forall \varepsilon \in (0,\infty),\, d\in \mathbb{N},\, \vp\in[0,\infty)^2, \tau >0\end{align*}

Observe that for this tractability function $T$, we have strong tractability if the following condition is satisfied
\[
 T(\varepsilon^{-1},1,\vp)= \phi(\varepsilon^{-1},p)\, \psi(1,0) = \phi(\varepsilon^{-1},p)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0,
\] where in this case $\vp$ is a scalar. Furthermore, we assume the following technical conditions. There exists a $K_{(p,q),\tau}$ depending on $p,q$ and $\tau$, but independent of $\varepsilon$ and $d$, such that
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau} &\le K_{(p,1),\tau} \phi (\varepsilon^{-1},\tau p)\\
      [\psi(d,q)]^{\tau} &\le K_{(1,q),\tau} \psi (d,\tau q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals,\, p,q\geq 0,\, \tau >0.
\end{align*}
Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strongly separable tractability
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation*}
Theorem \ref{thm_main_tract2} yields the following equivalent condition for 
separable tractability
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = L_{(p,q)\psi(d,q)}}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\]
%As the name implies, the roles of $\varepsilon^{-1}$ and $d$ are separated. 
\end{example}
\begin{example}[Non-separable symmetric tractability]
    Let the tractability function $T$, be defined by
\[
 T(\varepsilon^{-1},d,(p,q))= \exp\{\phi(\varepsilon^{-1},p)\, \psi(d,q)\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p\ge 0.
\]  with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$. Furthermore, we assume the following technical conditions. There exists a $K_{(p,q),\tau}$ depending on $p,q$ and $\tau$, but independent of $\varepsilon$ and $d$, such that
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau} &\le K_{(p,1),\tau} \phi (\varepsilon^{-1},\tau p)\\
      [\psi(d,q)]^{\tau} &\le K_{(1,q),\tau} \psi (d,\tau q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals,\, p,q\geq 0,\, \tau >0.
\end{align*}
\end{example}
}

@@

\section{Tractability on a Restricted Domain}

Suppose that the domain of interesting $(\varepsilon^{-1},d)$ is not all of $(0,\infty \times \naturals$ but given by 
\begin{equation}
    \Omega  = \{(\varepsilon^{-1},d) : \varepsilon > E(d) \ge 0, \ d \in \naturals \},
\end{equation}
where $E$ is a non-increasing function.  A problem is
$T$-tractable with parameter $\vp$ \emph{on the restricted domain $\Omega$} iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:rest_strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall(\varepsilon^{-1},d) \in \Omega.
\end{equation}
Define $M(d) = \max\{n : \lambda_{n+1,d} \ge E(d)\}$, i.e., 
\[
\lambda_{1,d}\ge  \cdots \ge \lambda_{M(d)+1,d} \ge E(d) > \lambda_{M(d)+2,d}, \qquad
(\lambda_{1,d}^{-1},d), \cdots, (\lambda^{-1}_{M(d)+1,d},d) \} \in \Omega.
\]

\begin{theorem}\label{thm_main_rest_tract2}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is $T$-tractable on $\Omega$ iff there exists $\vp\ge \vzero$ and a positive integer $L_{\vp}$ such that
\begin{equation} \label{eq:rest_tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^{M(d)} \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:rest_tractiff4} holds for some $\vp$, let $\widetilde{\cp}_{\textup{trct}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty}) \}$.
Then $\cp_{\textup{trct}} = \widetilde{\cp}_{\textup{trct}}$, and the set of optimal  tractability parameters is
\[
\cp_{\textup{opt}} =
\{\vp^* \in \widetilde{\cp}_{\textup{trct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{trct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

\begin{proof}
    \textbf{Sufficient Condition:}\\
We adapt the argument in the proof of the sufficient condition for Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}.  Suppose that \eqref{eq:rest_tractiff4} holds for some $\vp \ge 0$. Fix any $(\varepsilon^{-1},d) \in \Omega$.  Then

Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},d,\vp)$ are non-decreasing in $i$ for $i \in \{1, \ldots, M(d)+1\}$. In particular, for  any $n \in \{N, \ldots, M(d)\}$
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{N,d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},d,\vp)} \le \cdots \le \frac{1}{T(\lambda_{N, d}^{-1},d,\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d,\vp) }
    \le &\frac{1}{n-N+1} \sum_{i=N}^n  \frac{1}{T(\lambda_{i, d}^{-1},d,\vp) }
    \le \frac{1}{n-N+1} \sum_{i=N}^{M(d)}  \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}.
\end{align*}
Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a_rest}
    n - N +1 \ge T(\varepsilon^{-1},d,\vp) \sum_{i=N}^{M(d)} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},d,\vp)} \le
   \frac{1}{n-N+1} \sum_{i=N}^{M(d)} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp) } \le \frac{1}{T(\varepsilon^{-1},d,\vp)} \\ \forall d, n, N \in \naturals.
\end{multline}
We note that the left side of the implication implies $n \ge N$ automatically since $n$ is an integer  $n - N +1$ is assumed to be positive.

Moreover, given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity via \eqref{eq:comp_def}:
\begin{align} %\label{eq:comp_equiv}
	\nonumber
	\comp(\varepsilon, d) & = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
	\nonumber
	& = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},d,\vp)}\le \frac{1}{T(\varepsilon^{-1},d,\vp) }\biggr\}\\
	\nonumber
	& \le  \min \biggl\{n \in \natzero : n \ge N - 1 + T(\varepsilon^{-1},d,\vp) \sum_{i=N}^{M(d)} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)} \biggr\} \;  \forall N \in \naturals, \quad \text{by \eqref{eq:implicate_a_rest}} \\ 
	& \le  T(\varepsilon^{-1},d,\vp) \Biggl [\frac{N}{T(\varepsilon^{-1},d,\vp)}  +  \sum_{i=N}^{M(d)} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)} \Biggr] \quad \forall N \in \naturals. \label{eq:compUP_rest}
\end{align}

it follows from \eqref{eq:compUP} with $N = \lceil L_{\vp}\, T(0,d,\vp)\rceil$ that 
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [ 
       \frac{\lceil L_{\vp}\, T(0,d,\vp)\rceil}{T(\varepsilon^{-1},d,\vp)}  + 
       \underbrace{\sum_{i=\lceil L_{\vp}\, T(0,d,\vp)\rceil}^{M(d)} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:rest_tractiff4}}}
       \Biggr ] 
        \quad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)+1}{T(\varepsilon^{-1},d, \vp)} +  S_{\vp} \right]
        \qquad \forall \varepsilon > 0, \ d  \in \naturals\\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp} T(0,d,\vp)}{T(\varepsilon^{-1},d, \vp)}
       + \frac{1}{T(0,1, \vp)} + S_{\vp}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}   \\
       & \le T(\varepsilon^{-1},d,\vp) 
       \underbrace{\left [  L_{\vp} + \frac{1}{T(0,1, \vp)} + S_{\vp} \right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}  
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:rest_tractiff4}.
